---
title: "Linear Regression" 
author: "Yifei Sun"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
--- 

\newpage

In this example, we assess the association between high density lipoprotein (HDL) cholesterol and body mass index, blood pressure, and other demographic factors (age, gender, race) using the NHANES data (https://wwwn.cdc.gov/nchs/nhanes/ContinuousNhanes/Default.aspx?BeginYear=2001).
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = 'asis')
```


```{r}
library(RNHANES)
library(tidyverse)
library(summarytools)
library(stargazer)
library(leaps)
```

The data can be downloaded using functions in the package `RNHANES`.

```{r}
dat <- nhanes_load_data("l13_B", "2001-2002")

dat = dat %>% 
  left_join(nhanes_load_data("BMX_B", "2001-2002"), by="SEQN") %>%
  left_join(nhanes_load_data("BPX_B", "2001-2002"), by="SEQN") %>%
  left_join(nhanes_load_data("DEMO_B", "2001-2002"), by="SEQN")

dat = dat %>% 
  select(SEQN, RIAGENDR, RIDRETH1, RIDAGEYR, BMXBMI, BPXSY1, LBDHDL)

colnames(dat) <- c("ID", "gender", "race", "age", "bmi", "sbp", "hdl")

dat$race <- as.factor(dat$race)

dat <- na.omit(dat)
```

We first look at the summary statistics of the predictors and the response.

```{r}
st_options(plain.ascii = FALSE,       
           style = "rmarkdown", 
           dfSummary.silent = TRUE,        
           footnote = NA,          
           subtitle.emphasis = FALSE)      

dfSummary(dat[,-1])
```

# Simple linear regression

```{r, fig.height = 3, fig.width = 4, fig.align = 'center'}
fit0 <- lm(hdl ~ bmi,
           data = dat)

stargazer(fit0, header=FALSE, type='latex')

ggplot(dat, aes(bmi, hdl)) +
  geom_point(alpha = 0.05, size = 0.5, color = "black") +
  geom_smooth(method = "lm") + 
  labs(x = "BMI", y = "HDL cholesterol")
```


# Multiple linear regression: a small experiment

```{r}
fit1 <- lm(hdl ~ bmi + age + gender + race + sbp,
           data = dat)

fit2 <- lm(bmi ~ age + gender + race + sbp,
           data = dat)

r2 <- fit2$residuals

fit3 <- lm(hdl ~ r2,
           data = dat)

coef(fit1)["bmi"]

coef(fit3)["r2"]
```

# Prediction interval vs. confidence interval

```{r}
newdata <- dat[1,]
predict(fit1, newdata, interval = "predict") 
predict(fit1, newdata, interval = "confidence") 
```

# Model selection

```{r}
regsubsetsObj <- regsubsets(hdl ~ bmi + age + gender + race + sbp, data = dat, 
                            method = "exhaustive") 
plot(regsubsetsObj, scale = "adjr2")
```
